Supplement to: Learning to Infer Graphics Programs from Hand-Drawn Images

### 2018-04-25 revision.tex
nips 2018

\begin{abstract}
  We introduce a model that learns to convert simple hand drawings
  into graphics programs written in a subset of \LaTeX.~The model
  combines techniques from deep learning and program synthesis.  We
  learn a convolutional neural network that proposes plausible drawing
  primitives that explain an image. These drawing primitives are
  like a trace of the set of primitive commands issued by a graphics
  program. We learn a model that uses program synthesis techniques to
  recover a graphics program from that trace. These programs have constructs like variable
  bindings, iterative loops, or simple kinds of conditionals. With a
  graphics program in hand, we can correct errors made by the deep
  network %, measure similarity between drawings by use of similar high-level geometric
  %structures,
  and extrapolate drawings.  Taken together these results
  are a step towards agents that induce useful, human-readable
  programs from perceptual input.
\end{abstract}

@@

Even with the high-level idea of a trace set, going from hand drawings to programs remains difficult. We address  these challenges:
(1) Inferring trace sets from images requires \emph{domain-specific} design choices from the deep learning and computer vision toolkits (Sec.~\ref{neuralNetworkSection}). (2) Generalizing to noisy hand drawings requires learning a \emph{domain-specific} noise model that is invariant to the variations across hand drawings (Sec.~\ref{generalizingTheHandDrawings}).  (3) Discovering good programs requires solving a difficult combinatorial search problem, because the programs are often long and complicated (e.g., 9 lines of code, with nested loops and conditionals). We give a \emph{domain-general} framework for learning a search policy that quickly guides program synthesizers toward the target programs (Sec.~\ref{learningASearchPolicy}).

@@

  \textbf{Experiment 2.}
    We evaluated, but did not train, our system on 100 real hand-drawn figures; see Fig.~\ref{drawingSuccesses}--\ref{drawingFailures}.
These were drawn carefully but not perfectly with the aid of graph paper.
%Because our model assumes that objects are snapped to a $16\times 16$ grid, 
%we made the drawings on graph paper.
For each drawing we annotated a ground truth trace set and had the neurally guided SMC sampler
produce $10^3$ samples for each drawing.
For 63\% of the drawings, the Top-1 most likely sample exactly matches the
ground truth; with more samples, the model finds trace sets
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
We will show that the program synthesizer
corrects some of these small errors (Sec.~\ref{synthesizerHelpsParsing}).
%% Because the model sometimes makes mistakes % on hand drawings,
%% we envision it working as follows:
%% a user sketches a diagram,
%% and the system responds by proposing a few candidate interpretations.
%% The user could then select the one closest to their intention and edit it if necessary.

@@

\begin{figure}[h]\centering
  \begin{minipage}[c]{0.57\textwidth} 
    \centering  \includegraphics[width = 8cm]{figures/drawingAccuracy.png}            \vspace{-0.5cm}
  \end{minipage}\hfill%
      \begin{minipage}[c]{0.4\textwidth} 
  \caption{How close are the model's outputs to the ground truth on hand drawings, as we consider larger sets of samples (1, 5, 100)?
    Distance to ground truth trace set measured by the intersection over union of predicted vs. ground truth. (a) for 63\% of drawings the model's top prediction is exactly correct; (b) for 70\% of drawings the ground truth is in the top 5 model predictions; (c) for 4\% of drawings all of the model outputs have no overlap with the ground truth. Red: the full model. Other colors: lesioned versions of our model.}\label{drawingIntersectionOverUnion}            \vspace{-0.5cm}
      \end{minipage}

>>>>

\begin{abstract}
  We introduce a model that learns to convert simple hand drawings
  into graphics programs written in a subset of \LaTeX.~The model
  combines techniques from deep learning and program synthesis.  We
  learn a convolutional neural network that proposes plausible drawing
  primitives that explain an image. These drawing primitives are like
  a trace of the set of primitive commands issued by a graphics
  program. We learn a model that uses program synthesis techniques to
  recover a graphics program from that trace. These programs have
  constructs like variable bindings, iterative loops, or simple kinds
  of conditionals. With a graphics program in hand, we can correct
  errors made by the deep network and extrapolate drawings.  Taken
  together these results are a step towards agents that induce useful,
  human-readable programs from perceptual input.
\end{abstract}

@@

Even with the high-level idea of a trace set, going from hand drawings to programs remains difficult. We address  these challenges:
(1) Inferring trace sets from images requires \emph{domain-specific} design choices from the deep learning and computer vision toolkits (Sec.~\ref{neuralNetworkSection}). (2) Generalizing to noisy hand drawings, we will show, requires learning a \emph{domain-specific} noise model that is invariant to the variations across hand drawings (Sec.~\ref{generalizingTheHandDrawings}).  (3) Discovering good programs requires solving a difficult combinatorial search problem, because the programs are often long and complicated (e.g., 9 lines of code, with nested loops and conditionals). We give a \emph{domain-general} framework for learning a search policy that quickly guides program synthesizers toward the target programs (Sec.~\ref{learningASearchPolicy}).

@@

  \textbf{Experiment 2: Figures~\ref{drawingSuccesses}--\ref{drawingIntersectionOverUnion}.}
    We evaluated, but did not train, our system on 100 real hand-drawn figures; see Fig.~\ref{drawingSuccesses}--\ref{drawingFailures}.
These were drawn carefully but not perfectly with the aid of graph paper.
%Because our model assumes that objects are snapped to a $16\times 16$ grid, 
%we made the drawings on graph paper.
For each drawing we annotated a ground truth trace set and had the neurally guided SMC sampler
produce $10^3$ samples. % for each drawing.
For 63\% of the drawings, the Top-1 most likely sample exactly matches the
ground truth; with more samples, the model finds trace sets
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
We will show that the program synthesizer
corrects some of these small errors (Sec.~\ref{synthesizerHelpsParsing}).
%% Because the model sometimes makes mistakes % on hand drawings,
%% we envision it working as follows:
%% a user sketches a diagram,
%% and the system responds by proposing a few candidate interpretations.
%% The user could then select the one closest to their intention and edit it if necessary.

@@

\begin{figure}[h]\centering
  \begin{minipage}[c]{0.57\textwidth} 
    \centering  \includegraphics[width = 8cm]{figures/drawingAccuracy.png}            \vspace{-0.5cm}
  \end{minipage}\hfill%
      \begin{minipage}[c]{0.4\textwidth} 
  \caption{How close are the model's outputs to the ground truth on hand drawings, as we consider larger sets of samples (1, 5, 100)?
    Distance to ground truth trace set measured by the intersection over union (IoU) of predicted vs. ground truth: IoU of sets $A$ and $B$ is $|A\cap B|/|A\cup B|$. (a) for 63\% of drawings the model's top prediction is exactly correct; (b) for 70\% of drawings the ground truth is in the top 5 model predictions; (c) for 4\% of drawings all of the model outputs have no overlap with the ground truth. Red: the full model. Other colors: lesioned versions of our model.}\label{drawingIntersectionOverUnion}            \vspace{-0.5cm}
      \end{minipage}

### 2018-04-25 supplement.tex
nips 2018

@@

\subsection{Convolutional network}
The convolutional network takes as input 2 $256\times 256$ images
represented as a $2\times 256\times 256$ volume. These are
passed through two layers of convolutions separated by ReLU
nonlinearities and max pooling:
\begin{itemize}
\item Layer 1: 20 $8\times 8$ convolutions, 2 $16\times 4$ convolutions, 2 $4\times 16$ convolutions. Followed by $8\times 8$ pooling with a stride size of 4.
\item Layer 2: 10 $8\times 8$ convolutions. Followed by $4\times 4$ pooling with a stride size of 4.
\end{itemize}

@@

\subsection{Predicting continuous coordinates using Mixture Density Networks}

@@

\begin{figure}
  \includegraphics[width = \textwidth]{figures/continuousParses.png}
  \caption{Parsing hand drawings using the MDN variant of our
    model. Black-on-white: hand drawings. White-on-black, below each
    hand drawing: the parse inferred by the model.}\label{continuousParses} 
  \end{figure}
\subsection{LSTM Baseline}

>>>>

\section{Further testing of the trace hypothesis}

Recall that we factored the graphics program synthesis problem into two components: (1) a perception-facing component, whose job is to go from perceptual input to a set of commands that must occur in the execution trace of the program (\textbf{trace set}); and (2) a program synthesis component, whose job is to infer a program whose execution trace contains those commands. This is a different approach from other recent program induction models (e.g.,~\cite{devlin2017robustfill,nps}), which regress directly from a program induction problem to the source code of the program. We motivated the usefulness of our factoring of the problem via The Trace Hypothesis.

\textbf{Experiment.} To test whether the trace hypothesis is necessary for our domain, we trained a model to regress directly from images to graphics programs. This baseline model, which we call the \emph{no-trace baseline}, was able to infer some simple programs, but failed completely on more sophisticated scenes.

Baseline model architecture: The model architecture is a straightforward, image-captioning-style CNN$\to$LSTM. We keep the same CNN architecture from our main model (Section~\ref{CNNarchitecture}), with the sole difference that it takes only one image as input. The LSTM decoder produces the program token-by-token: so we flatten the program's hierarchical structure, and use special ``bracketing'' symbols to convey nesting structure, in the spirit of~\citep{vinyals2015grammar}. The LSTM decoder has 2 hidden layers with 1024 units. We used 64-dimensional embeddings for the program tokens.

Training and evaluation: The model was trained on $10^7$ synthetically generated programs -- 2 orders of magnitude more data than the model we present in the main paper. We then evaluated the baseline on \emph{synthetic renders} of our 100 hand drawings (the testing set used throughout the paper). Recall that our model was evaluated on noisy real hand drawings. We sample programs from this baseline model conditioned on a synthetic render of a hand drawing, and report only the sampled program whose execution trace most closely matched the ground truth trace set, as measured by the symmetric difference of the two sets. We allow the baseline model to spend 1 hour drawing samples per drawing -- recall that our model finds 58\% of programs in under a minute. Together these training and evaluation choices are intended to make the problem as easy as possible for the no-trace baseline.

Results: The no-trace  baseline succeeds for trivial programs (a few lines, no variables, loops, etc.); occasionally gets small amounts of simple looping structure; and fails utterly for most of our test cases. See Figure~\ref{noTrace}.

\begin{figure}[h]\centering
      \includegraphics[width = 2cm]{figures/noTraceOutputs/11.png}
  \includegraphics[width = 2cm]{figures/noTraceOutputs/16.png}
    \includegraphics[width = 2cm]{figures/noTraceOutputs/29.png}
    \includegraphics[width = 2cm]{figures/noTraceOutputs/7.png}
        \includegraphics[width = 2cm]{figures/noTraceOutputs/94.png}
      \includegraphics[width = 2cm]{figures/noTraceOutputs/38.png}
  \caption{Top, white: synthetic rendering of a hand drawing. Bottom, black: output of best program found by no-trace baseline.}\label{noTrace}
  \end{figure}

@@

\subsection{Convolutional network}\label{CNNarchitecture}
The convolutional network takes as input 2 $256\times 256$ images
represented as a $2\times 256\times 256$ volume. These are
passed through two layers of convolutions separated by ReLU
nonlinearities and max pooling:
\begin{itemize}
\item Layer 1: 20 $8\times 8$ convolutions, 2 $16\times 4$ convolutions, 2 $4\times 16$ convolutions. Followed by $8\times 8$ pooling with a stride size of 4.
\item Layer 2: 10 $8\times 8$ convolutions. Followed by $4\times 4$ pooling with a stride size of 4.
\end{itemize}

@@

\begin{comment}
\subsection{Predicting continuous coordinates using Mixture Density Networks}

@@

\begin{figure}
  \includegraphics[width = \textwidth]{figures/continuousParses.png}
  \caption{Parsing hand drawings using the MDN variant of our
    model. Black-on-white: hand drawings. White-on-black, below each
    hand drawing: the parse inferred by the model.}\label{continuousParses} 
  \end{figure}
\end{comment}
\subsection{LSTM Baseline}
